---
title: "CODICEPERSONA_PROBLEMA"
author: "Marco Scarpelli"
date: "DATA"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
graphics.off()

library(MASS)
library(car)
library(rgl)   #3D plots
library(leaps) #best subset selection
library(tree)  #decision trees
library(corrplot) #correlation
library(glmnet)

#data <- read.table('NOME FILE', header=T)
```
# Print dataframe
```{r echo = FALSE} 
# Questa cella viene eseguita e viene mostrato il suo output ma non il codice.

df<-read.table('landslides.txt', header=T)
head(df)
dim(df)

```
# Point a
```{r echo = FALSE} 
fm <- lm(rate  ~ rain + hardness + coarse + fine
, data=df) #+ ... + regressor(r+1)) 
summary(fm)
coefficients(fm)
```
Our assumptions are that the residuals have 0 mean and are homoscedastic. Let us check for their Gaussianity:
```{r echo = FALSE} 
shapiro.test(residuals(fm))

par(mfrow=c(2,2))
plot(fm)
par(mfrow=c(1,1))
```

Furthermore, we can see that the Q-Q plot follows the line closely enough, and all points on the residual-leverage plot are within Cook's distance.

Let us also check the variance inflation factor:
```{r echo = FALSE} 
vif(fm)
```
where we can see that all parameters are well below 5 and especially 10.

# Point b
From the summary, we can see that the variable `hardness` is strongly not significant for our model. We will remove it:
```{r echo = FALSE} 
fm2 <- lm(rate  ~ rain + coarse + fine
, data=df) #+ ... + regressor(r+1)) 
summary(fm2)
coefficients(fm2)
```
where we can see that now all parameters seem to be significant for our model. Let us check whether the two models are equivalent:
```{r echo = FALSE} 
anova(fm,fm2)
```
According to the output's p-value (0.81), the two models perform equal predictions.

# Point c
```{r echo = FALSE} 
C=rbind(c(0,0,1,-2))
hyp0<-c(0)

linearHypothesis(fm2, C, hyp0) 
```
The null hypothesis is thus proven; we can update the model dataframe with a new column to account for this.

## Attempt 1
> Dubbio: qui secondo me avrebbe senso creare un nuovo dato nel seguente modo ed usare questo nuovo modello; l'altra soluzione è togliere coarse completamente e lasciare solo fine, ma secondo me non ha troppo senso.

The new column will contain the sum of coarse and 2$\times$fine.

```{r echo = FALSE} 
#df$coarse_new <- 2*df$fine + df$coarse
fm3 <- lm(rate  ~ rain + coarse + fine, data=df)
summary(fm3)
coefficients(fm3)
```


Let us check this w.r.t. the other models:
```{r echo = FALSE} 
anova(fm,fm2,fm3)
```

This new model is a bit different from the others.

## Attempt 2
We will remove `coarse`.
```{r echo = FALSE} 
#df$coarse_new <- 2*df$fine + df$coarse
fm3.2 <- lm(rate  ~ rain + fine, data=df)
summary(fm3.2)
coefficients(fm3.2)
```


Let us check this w.r.t. the other models:
```{r echo = FALSE} 
anova(fm,fm2,fm3.2)
```

This new model is a bit different from the others.

# Point d
```{r echo = FALSE} 
Z0.new <- data.frame(rain=700, hardness=5, coarse=10, fine=8)
alpha.pred <- 0.99
Conf <- predict(fm3.2, Z0.new, 
                interval='confidence', level=alpha.pred)  
t(Conf) # fit is the first value!

```